Minimizing post shock forecasting error using disparate information

### 2020-06-07 manuscript/manuscript.tex
updated abstract and edits to RMSE

Things to do:
\begin{enumerate}
  \item Literature review and introduction. Need more references, need to mention synthetic interventions in the introduction or the discussion.
  \item Need to mention possible generalizations in the Discussion.
  \item Need to add weighted adjustment estimator to the data analysis.
  \item Need to add simulations. 
  \item Update abstract.
\end{enumerate}

@@

\begin{enumerate}
  \item Remark for non-uniqueness of $\mathbf{W}^*$ is added in Section \ref{properties}.
  \item Generalizations are added in Section 6
  \item Data analysis is updated with results of weighted adjustment estimators
  \item Some preliminary setup of the simulation is added in Section 4.
\end{enumerate}

@@

\maketitle
\begin{abstract}
    We develop a forecasting methodology for time series data that has 
    undergone a shock. We still can provide credible forecasts for a time 
    series in the presence of such systematic shocks by drawing from disparate 
    time series that have undergone similar shocks for which post-shock 
    outcome data is recorded.  These disparate time series are assumed to have 
    mechanistic similarities to the time series under study but are otherwise 
    independent (Granger noncausal).  The inferential goal of our forecasting 
    methodology is to supplement observed time series data with post-shock 
    data from the disparate time series in order to minimize average forecast 
    risk. 
\end{abstract}

@@

We will suppose that an analyst has time series data ($y_{i,t}$,$\x_{i,t}$), 
$t = 1$, $\ldots$, $T_i$, $i = 1$, $\ldots$, $n+1$, where $y_{i,t}$ is a 
scalar response and $\x_{i,t}$ is a vector of covariates that are revealed to 
the analyst prior to the observation of $y_{1,t}$.  Suppose that the analyst 
is interested in forecasting $y_{1,t}$, the first time series in the 
collection.
%$y_1, y_2, \ldots$.  
%Given each time point $t \geq 1$, let $\x_{1,t}$ be a vector of covariates 
%revealed prior to the observation of $y_{1,t}$.  
To gauge the performance of a procedure that produces forecasts 
$\{\hat y_{1,t}, t= 1,2,\ldots\}$ given time horizon $T_1$, 
we consider forecast risk in the form of root mean squared error (RMSE),
$$
  R_T = \sqrt{\frac{1}{T}\sum_{t=1}^T\E{\hat y_{1,t} - y_{1,t}}^2},
$$
in our analyses. \textbf{Let's switch to root mean squared error, this metric is often used in forecasting articles.} In this article, we consider a dynamic panel data 
model with autoregressive structure similar to that in \citet{blundell1998initial}. 
%without past time period covariate 
%information and with a single present time covariate.  
Our dynamic panel model includes an additional shock effect whose presence 
or absence is given by the binary variable $D_{i,t}$, the details of this model 
are in the next section.

@@

\begin{enumerate}
\item[(1)] {\bf AR(1) model and time window}. We will use a simple AR(1) model to forecast Conoco Phillips stock price. This model has been shown to beat no-change forecasts when predicting oil prices over time horizons of 1 and 3 months \citep{alquist2013forecasting}. We will consider 30 pre-shock trading days and we will forecast the immediate shock effect and the shock effect over a future five trading day window. All estimates will be adjusted for inflation. The model setup for AR(1) is exactly the same as what is stated in Section \ref{modelsetup} with addition of shock effects. All the parameters are estimated using OLS.
\item[(2)] {\bf Selection of covariates}. We perform our analyses incorporating daily S\&P 500 index prices and West Texas Intermediate (WTI) crude oil prices as covariates. 
\item[(3)] {\bf Construction of donor pool}. Our donor pool will consist of Conoco Phillips shock effects observed after September 11th, 2001, %March 18th, 2008, 
several events in September, 2008, and November 27, 2014. The first two shock effects were observed during recessions that were predicated by an inversion of the yield curve \citep{bauer2018economic}, and the third was an OPEC induced supply side shock effect \citep{huppmann2015opec}. The reasons for those three shocks are:
  \begin{enumerate}
  \item On September 11th, 2001, Islamic extremist al-Qaeda committed a series of terrorist attacks against the United States of America \citep{braniff2011towards}. In addition to the tragic loss of life, this triggered negative repercussions on the U.S. and world economies with falling stock prices and a falling U.S. dollar \citep{floyd2001terror}. Moreover, trading was closed until September 17, 2001 \citep{johnston2006impact}.
  \item  In early September 2008, time series of oil prices experienced a sudden increase in volatility simultaneously due to turmoil in financial markets. The political, economic, social or environmental events may coincide with these shocks \citep{ewing2013volatility}. Notable shock effects followed the placement of Fannie May and Freddie Mac in conservatorship on September 7th (shock effect on the 8th), Lehman Brothers filing for bankruptcy on September 15th, and the Office of Thrift Supervision closes Washington Mutual Bank on September 25th \citep{dwyer2009financial, longstaff2010subprime}.
  \item On November 27th, 2014, it is documented that oil prices fall as OPEC opts not to cut production \citep{huppmann2015opec}. During the Great Recession when economic activity clearly declined, both oil and stock prices fell which points to demand factors. During the second half of 2014, oil prices plummeted but equity prices generally increased, suggesting that supply factors were the key driver \citep[Page 19]{baffes2015great}. 
  \end{enumerate}
%\item[(4)] {\bf Forecast}. \textbf{Jilei, please fill this in. I do not know what you have in mind.}
We assume that the five shocks are independent of the shock that Conoco Phillips  experienced on 2020-03-09.
\end{enumerate}

@@

$\mathbf{W}^*$ reports that the $\alpha_1$ is similar to $\alpha_3$ (2008--09--08) and $\alpha_5$ (2014--11--27) according to the setup of $\mc{M}_2$. That is, the shock that Conoco Phillips experienced in 2020-03-09 is similar to the shock induced by the placement of Fannie May and Freddie Mac, and the one caused by oil supply shock on 2014--11--27. It is consistent with discussion at the beginning of the section.
\begin{table}[H]
  \caption{Bootstrap estimates and results yielded by risk-reduction propositions with $B = 1000$}\label{table2}
  \begin{center}
    \begin{tabular}{lrrr}
      & $\hat{\alpha}_{\rm adj}$ & $\hat{\alpha}_{\rm wadj}$ & $\hat{\alpha}_{\rm IVW}$ \\
      \hline 
    Bootstrapped mean  & $-5.340$ & $-5.014$ & $-5.287$ \\
    Bootstrapped variance & 0.508 & 0.480 & 2.821 \\
    Risk-reduction proposition & 1 & 1 & 1 \\
    Best shock-effect estimator & 0 & 0 & 1
    \end{tabular}
  \end{center}  
\end{table}
\vspace{-.5cm}

Using the parametric bootstrap proposed in Section \ref{varbootstrap}, we estimated parameters for risk-reduction propositions proposed in  Section \ref{properties}; and the result is presented in Table \ref{table2}. Note that this parametric bootstrap procedure is conditioned  on the  donor pool. Table \ref{table2} tells that all three proposed  shock-effect estimators are usable in the sense that the forecast employing them is better than the one without; and the $\hat{\alpha}_{\rm wadj}$ is the best in terms of the estimated risk-reduction quantity. Then, we verify the consistency of the result yielded by risk-reduction propositions with the reality as below.

@@

We can see from Figure~\ref{Fig:CP} that $\hat{\alpha}_{\text{adj}}$, $\hat{\alpha}_{\text{wadj}}$ and $\hat{\alpha}_{\text{IVW}}$ perform decently well, they do not recover the magnitude of the shock effect but are much better than unadjusted forecasts that do not account for shock effects. 
The unadjusted forecast has an RMSE of 9.870 whereas 
the use of $\hat{\alpha}_{\text{adj}}$, $\hat{\alpha}_{\text{adj}}$, and $\hat{\alpha}_{\text{IVW}}$ have  RMSE of 4.475,  4.432, and 4.741, respectively.
Therefore, results proposed by Table \ref{table2} are verified with the reduced risks for forecasts using  $\hat{\alpha}_{\text{adj}}$, $\hat{\alpha}_{\text{adj}}$, and $\hat{\alpha}_{\text{IVW}}$ than without, and that $\hat{\alpha}_{\rm wadj}$ is the best with lowest RMSE.

We also study additive shock effect estimators where the shock effects corresponding to separate supply and demand shocks are added to estimate the unknown shock effect. The supply shock donor pool consists of the November 27th, 2014 shock effect and the demand shock donor pool consists of the remaining shock effects. The additive adjustment estimator computed by adding the $\hat{\alpha}_{\text{adj}}$, $\hat{\alpha}_{\text{adj}}$, and $\hat{\alpha}_{\text{IVW}}$ estimators for the demand and supply shock effects without September 11th have  RMSEs of $1.341$, $1.677$, and $0.661$, respectively.

@@

@@

Our proposed model in Section \ref{modelsetup} is a simple model with results in Section \ref{properties} easily generalizing to other more complicated models. First, the model can be extended to AR($p$) for $y_{i,t}$ with an additional component $\alpha_i 1(D_{i,t}=T_i^* + 1)$.  The model for $\alpha_i$ in $\mc{M}_2$ may be generalized to $\alpha_i = \mu_{\alpha} + \delta_i' \mathbf{x}_{i, T_i^*+1} + \cdots + \gamma_i \mathbf{x}_{i, T_i^* -(p+1)} + \tilde{\varepsilon}_{i}$ with some proper simple adjustments for SCM method. Second, multiple shock-effects can be nested within a time series; and time series in the donor pool are allowed be dependent. This generalization was adopted in Section \ref{forecasting}. In this case, the variance expressions proposed in Section \ref{properties} will not work. However, the risk-reduction propositions in Section \ref{properties} will work  as long as those shock-effects (not necessarily mutually independent) in the donor pool are independent of the one of interest. Third, though it is reflected in $\mc{M}_2$, we stress that our proposed methods allow $\alpha_i$ to follow arbitrary distributions with existing first and second moments. Fourth, the covariates in the model for $\alpha_i$ in $\mc{M}_2$ can be different from the covariates in the model of $y_{i,t}$. It is because in the derivation such dependence is not used.

>>>>

Things to do:
\begin{enumerate}
  \item Literature review and introduction. Need more references, need to mention synthetic interventions in the introduction or the discussion.
  \item Need to add simulations. 
\end{enumerate}

@@

\begin{enumerate}
  \item Remark for non-uniqueness of $\mathbf{W}^*$ is added in Section \ref{properties}.
  \item Generalizations are added in Section 6
  \item Data analysis is updated with results of weighted adjustment estimators
  \item Some preliminary setup of the simulation is added in Section 4.
  \item RMSE is changed to MSE.
  \item Abstract is updated
\end{enumerate}

@@

\maketitle
\begin{abstract}
    We developed a forecasting methodology for providing credible forecasts for time series data that has undergone a shock  by borrowing knowledge from disparate 
    time series that have undergone similar shocks for which post-shock 
    outcome is recorded. Three shock-effects estimators were constructed for minimizing average forecast risk. We proposed conditions when our methodology works, and which estimator is the best in average forecast risk reduction. Parametric bootstrap procedures are provided for such realization. Several simulated data examples, and a real data example of forecasting Conoco Phillips stock price are provided for verification and illustration.
\end{abstract}

@@

We will suppose that an analyst has time series data ($y_{i,t}$,$\x_{i,t}$), 
$t = 1$, $\ldots$, $T_i$, $i = 1$, $\ldots$, $n+1$, where $y_{i,t}$ is a 
scalar response and $\x_{i,t}$ is a vector of covariates that are revealed to 
the analyst prior to the observation of $y_{1,t}$.  Suppose that the analyst 
is interested in forecasting $y_{1,t}$, the first time series in the 
collection.
%$y_1, y_2, \ldots$.  
%Given each time point $t \geq 1$, let $\x_{1,t}$ be a vector of covariates 
%revealed prior to the observation of $y_{1,t}$.  
To gauge the performance of a procedure that produces forecasts 
$\{\hat y_{1,t}, t= 1,2,\ldots\}$ given time horizon $T_1$, 
we consider forecast risk in the form of  mean squared error (MSE),
$$
  R_T = \frac{1}{T}\sum_{t=1}^T\E{\hat y_{1,t} - y_{1,t}}^2,
$$
in our analyses. In this article, we consider a dynamic panel data 
model with autoregressive structure similar to that in \citet{blundell1998initial}. 
%without past time period covariate 
%information and with a single present time covariate.  
Our dynamic panel model includes an additional shock effect whose presence 
or absence is given by the binary variable $D_{i,t}$, the details of this model 
are in the next section.

@@

\begin{enumerate}
\item[(1)] {\bf AR(1) model and time window}. We will use a simple AR(1) model to forecast Conoco Phillips stock price. This model has been shown to beat no-change forecasts when predicting oil prices over time horizons of 1 and 3 months \citep{alquist2013forecasting}. We will consider 30 pre-shock trading days and we will forecast the immediate shock effect and the shock effect over a future five trading day window. All estimates will be adjusted for inflation. The model setup for AR(1) is exactly the same as what is stated in Section \ref{modelsetup} with addition of shock effects. All the parameters are estimated using OLS.
\item[(2)] {\bf Selection of covariates}. We perform our analyses incorporating daily S\&P 500 index prices and West Texas Intermediate (WTI) crude oil prices as covariates. 
\item[(3)] {\bf Construction of donor pool}. Our donor pool will consist of Conoco Phillips shock effects observed after September 11th, 2001, %March 18th, 2008, 
several events in September, 2008, and November 27, 2014. The first two shock effects were observed during recessions that were predicated by an inversion of the yield curve \citep{bauer2018economic}, and the third was an OPEC induced supply side shock effect \citep{huppmann2015opec}. The reasons for those three shocks are:
  \begin{enumerate}
  \item On September 11th, 2001, Islamic extremist al-Qaeda committed a series of terrorist attacks against the United States of America \citep{braniff2011towards}. In addition to the tragic loss of life, this triggered negative repercussions on the U.S. and world economies with falling stock prices and a falling U.S. dollar \citep{floyd2001terror}. Moreover, trading was closed until September 17, 2001 \citep{johnston2006impact}.
  \item  In early September 2008, time series of oil prices experienced a sudden increase in volatility simultaneously due to turmoil in financial markets. The political, economic, social or environmental events may coincide with these shocks \citep{ewing2013volatility}. Notable shock effects followed the placement of Fannie May and Freddie Mac in conservatorship on September 7th (shock effect on the 8th), Lehman Brothers filing for bankruptcy on September 15th, and the Office of Thrift Supervision closes Washington Mutual Bank on September 25th \citep{dwyer2009financial, longstaff2010subprime}.
  \item On November 27th, 2014, it is documented that oil prices fall as OPEC opts not to cut production \citep{huppmann2015opec}. During the Great Recession when economic activity clearly declined, both oil and stock prices fell which points to demand factors. During the second half of 2014, oil prices plummeted but equity prices generally increased, suggesting that supply factors were the key driver \citep[Page 19]{baffes2015great}. 
  \end{enumerate}
%\item[(4)] {\bf Forecast}. \textbf{Jilei, please fill this in. I do not know what you have in mind.}
We assume that the five shocks are independent of the shock that Conoco Phillips  experienced on 2020-03-09. The covariates and response of time series in the donor pool are adjusted for inflation.
\end{enumerate}

@@

$\mathbf{W}^*$ reports that the $\alpha_1$ is similar to $\alpha_3$ (2008--09--08) and $\alpha_5$ (2014--11--27) according to the setup of $\mc{M}_2$. That is, the shock that Conoco Phillips experienced in 2020-03-09 is similar to the shock induced by the placement of Fannie May and Freddie Mac, and the one caused by oil supply shock on 2014--11--27. It is consistent with discussion at the beginning of the section.
\begin{table}[H]
  \caption{Bootstrap estimates and results yielded by risk-reduction propositions with $B = 1000$}\label{table2}
  \begin{center}
    \begin{tabular}{lrrr}
      & $\hat{\alpha}_{\rm adj}$ & $\hat{\alpha}_{\rm wadj}$ & $\hat{\alpha}_{\rm IVW}$ \\
      \hline 
    Bootstrapped mean  & $-5.340$ & $-5.014$ & $-5.287$ \\
    Bootstrapped variance & 0.508 & 0.480 & 2.821 \\
    Risk-reduction proposition & 1 & 1 & 1 \\
    Best shock-effect estimator & 0 & 1 & 0
    \end{tabular}
  \end{center}  
\end{table}
\vspace{-.5cm}

Using the parametric bootstrap proposed in Section \ref{varbootstrap}, we estimated parameters for risk-reduction propositions and risk-reduction quantities proposed in  Section \ref{properties}; and the result is presented in Table \ref{table2}. Note that this parametric bootstrap procedure is conditioned  on the  donor pool. Table \ref{table2} tells that all three proposed  shock-effect estimators are usable in the sense that the forecast employing them is better than the one without; and the $\hat{\alpha}_{\rm wadj}$ is the best in terms of the estimated risk-reduction quantity. Then, we verify the consistency of the result yielded by risk-reduction propositions with the reality as below.

@@

We can see from Figure~\ref{Fig:CP} that $\hat{\alpha}_{\text{adj}}$, $\hat{\alpha}_{\text{wadj}}$ and $\hat{\alpha}_{\text{IVW}}$ perform decently well, they do not recover the magnitude of the shock effect but are much better than unadjusted forecasts that do not account for shock effects. 
The unadjusted forecast has an MSE of 97.422 whereas 
the use of $\hat{\alpha}_{\text{adj}}$, $\hat{\alpha}_{\text{wadj}}$, and $\hat{\alpha}_{\text{IVW}}$ have  MSE of 20.028,  19.647, and 22.477, respectively.
Therefore, results proposed by Table \ref{table2} are verified with the reduced risks for forecasts using  $\hat{\alpha}_{\text{adj}}$, $\hat{\alpha}_{\text{wadj}}$, and $\hat{\alpha}_{\text{IVW}}$ than without, and that $\hat{\alpha}_{\rm wadj}$ is the best with lowest MSE.

We also study additive shock effect estimators where the shock effects corresponding to separate supply and demand shocks are added to estimate the unknown shock effect. The supply shock donor pool consists of the November 27th, 2014 shock effect and the demand shock donor pool consists of the remaining shock effects. The additive adjustment estimator computed by adding the $\hat{\alpha}_{\text{adj}}$, $\hat{\alpha}_{\text{adj}}$, and $\hat{\alpha}_{\text{IVW}}$ estimators for the demand and supply shock effects have  MSEs of $1.797$, $2.812$, and $0.437$, respectively.

@@

Our proposed model in Section \ref{modelsetup} is a simple model with results in Section \ref{properties} easily generalizing to other more complicated models. First, the model can be extended to AR($p$) for $y_{i,t}$ with an additional component $\alpha_i 1(D_{i,t}=T_i^* + 1)$.  The model for $\alpha_i$ in $\mc{M}_2$ may be generalized to $\alpha_i = \mu_{\alpha} + \delta_i' \mathbf{x}_{i, T_i^*+1} + \cdots + \gamma_i \mathbf{x}_{i, T_i^* -(p+1)} + \tilde{\varepsilon}_{i}$ with some proper simple adjustments for SCM method.

Second, multiple shock-effects can be nested within a time series; and time series in the donor pool are allowed be dependent. This generalization was adopted in Section \ref{forecasting}. In this case, the variance expressions proposed in Section \ref{properties} will not work. However, the risk-reduction propositions in Section \ref{properties} will work  as long as those shock-effects (not necessarily mutually independent) in the donor pool are independent of the one of interest.

Third, though it is reflected in $\mc{M}_2$, we stress that our proposed methods allow $\alpha_i$ to follow arbitrary distributions with existing first and second moments. Fourth, the covariates in the model for $\alpha_i$ in $\mc{M}_2$ can be different from the covariates in the model of $y_{i,t}$. It is because in the derivation such dependence is not used. Fifth, the metric used in construction of $\hat{\alpha}_{\rm wadj}$ may be replaced with other complete metrics.

@@
